\chapter{Quantum Error Correction}

For the following chapter I used multiple previous works as a reference, which I can recommend for anybody interested \cite{terhal_qec_formalism}, \cite{intro_to_qec} and \cite{gottesmann_book}.

Why do we bother to develop quantum computing?

Quantum computing is interesting because it offers the possibility to use quantum phenomena like superposition and entanglement in algorithms.
Such algorithms can solve some problems efficiently for which we do not yet know an efficient classical algorithm.
A famous examples is the problem of integer factorization, as proposed by Shor \cite{shor_factoring}. % because the problem of integer factorziation was the basis for a lot of encryption
Other examples are the grover search algorithm \cite{grover_search_algo} or this \cite{chemistry_example_algo} proposed algorithm to calculate the energy of atoms and molecules.
% Show the bitcoin paper for fun?

To realize quantum computations we first need a quantum state in which we can store our quantum information, analogous to the (classical) bit of a (classical) computer.
Such a state is called a quantum bit (qubit)\footnote{
    It is only called a qubit if it is a binary/two level systems. One can also realize a higher dimensional data states, so called qudits, but they are not the topic of this thesis and therefore not discussed further. %[TODO: name paper by Irina?]
} in the quantum case.
To implement the proper algorithms we also need to be able to do operations on the quantum information, but the implementation of those is out of scope of this thesis. 
We will focus on maintaining the quantum information contained in one qubit over time, a so called quantum memory experiment. 

There are many different approaches of realizing qubits as physical systems, 
like trapped ions,  % TODO find references
neutral ions,       % TODO find references
molecules,          % TODO find references
photons,            % TODO find references
superconducting circuits,% TODO find references
and certainly more. 
Each with their own advantages and disadvantages, but all of them still share the same problem they have a residual coupling with the environment and are therefore susceptible to faults. 
Those faults can introduce errors which could change the information contained in the qubit and therefore destroy our quantum information memory. 

The quantum information is stored in a single excited state of an ion, a single electron, for the example of trapped ions. 
In comparison the classical information contained in one bit is stored in approximately $10^4$ electrons \cite{classical_bits_electrons}.
Therefore one can intuitively understand that a qubit is intrinsically less robust, than a classical bit in regards to errors.
% An because we still need to stay in the quantum regime we can not simply increase the number of electrons for each physical qubit, this would destroy/decohere the quantum information. % is this really correct?
% This is the approach to make my mother understand why we need QEC.


To mitigate the influence of errors and safeguard the quantum information against errors, quantum error-correction (QEC) codes were developed. %should put same papers here
Quantum error-correction codes are different from classical error-correcting codes, because we have to take into account that by measuring we can destroy superpositions (Wavefunction collapse).
Moreover where there is only one type of error in classical computers, the bit flip, quantum computers can have a multitude of different errors, as will be discussed in section \ref{sec:pauli_errors}.

In QEC codes we take active measures to correct the quantum memory and remove errors from the qubits. 
This is done by encoding one\footnote{We can also encode more than one qubit at a time, but this is also not the focus of this work.} (logical) qubit in many (physical) qubits. 
Using the redundancy in those qubits we can extract some information, which does not tell us anything about the logical state, but can be used to infer which errors occurred.
And based on this information we can take actions to correct those occured errors \cite{terhal_qec_formalism}.

We will got into more detail in section \ref{sec:stabilizer_formalism}, where we focus on stabilizer codes, which are a class of quantum error-correcting codes and which we are gonna use in this thesis.
But before we can discuss how to correct errors we first need to define what our errors are.

% should I include a more general defintion of QECC and an abstract circuit example or equations, to explain what exactly defines a QECC.
% namely encoding + information extraction + decoding + correction

% introduce pauli errors here?

\section{Stabilizer Codes and their Structure}\label{sec:stabilizer_formalism}

One class of QEC codes are the stabilizer codes \cite{gottesmann_book},\cite{gottesmann_thesis}.
This section is mainly based by Gottesmann works, as cited above, and the work of Terhal \cite{terhal_qec_formalism}.

In this and in all the following chapters we limit ourselves to Pauli operators as possible errors. 
The argumentation why this is assumption valid is following in section \ref{sec:pauli_errors}.

The goal of this chapter is to introduce the structure and formalism behind stabilizer codes.
We want to place a special focus on the group structure behind stabilizer codes to motivate the later chapter about possible decoder approaches. 

\subsection{Stabilizer Group}

The stabilizer formalism is based around a subgroup of the Pauli group $\mathcal{P}_n$ called the stabilizer group $\mathcal{S}$. 
The reason for the name becomes obvious in the next subsection \ref{ssec:codespace}, when we talk about the codespace.

By choosing a certain set of Pauli operator one defines the stabilizer code.
We will later in section \ref{sec:kitaev_surface_code} discuss a specific choice but we first take a general view upon stabilizer codes. 
One can freely choose which operators to use to define a stabilizer group with the following limitations for the resulting stabilizer group.
It must be an Abelian subgroup of the $n$-qubit Pauli group $\mathcal{P}_n$ and it is not allowed to contain $-I$. 

Many of the following definitions will be formulated by only taking a set of $m$ linear independent generators of the stabilizer group, which we will label as $S_i$, such that the stabilizer group can be written as $\langle S_1, S_2, ..., S_m\rangle= \mathcal{S}$.
To shorten the notation even further we will refer to the set of linear independent generators of the stabilizer group by $\mathcal{S}_G$.
We will also refer to the linear independent generators of the stabilizer group by generators to shorten the following text.


\subsection{Codespace}\label{ssec:codespace}% rework this headline 

As described earlier the basic idea behind QEC codes is to encode the logical information of $k$ logical qubits into $n$ physical qubits.  
The space in which the logical states exist is called the codespace $\mathcal{L}$, which is a subspace of the full $n$-qubit Hilbert space $\mathcal{H}_{2^n}$. 

Using the stabilizer group defined in the previous subsection, we can define the codespace $\mathcal{L}$ as being the space spanned by all states $\ket{\Psi}$ which are invariant under all elements of the stabilizer group.

\begin{equation}
    \mathcal{L} = \left\{  \ket{\Psi} \in  \mathcal{H}_{2^n} : S \ket{\Psi} = \ket{\Psi} \forall S \in \mathcal{S} \right\}
    \label{eq:defintion_codespace_whole_group}
\end{equation}

Given that each element of the stabilizer can be generated by our set of generators $\mathcal{S}_G$, we can rewrite this definition as 

\begin{equation}
    \mathcal{L} = \left\{  \ket{\Psi} \in  \mathcal{H}_{2^n} : S \ket{\Psi} = \ket{\Psi} \forall S_i \in \mathcal{S}_G \right\}.
    \label{eq:defintion_codespace_generators}
\end{equation}

Given that we have $n$ physical qubits into which we encode the logical information and used a stabilizer group with $m$ linear independent generators, we can encode $k=n-m$ logical qubits into the resulting codespace. 
The resulting codespace is therefore a $2^k$ dimensional Hilbert space.

We will refer to the states which are in the codespace, as codewords $\ket{\Phi}\in \mathcal{L}$.

Here we note explicitly that acting with an element of the stabilizer group on any codeword is equivalent to act with the identity
\begin{equation}
    S \ket{\Phi} = \ket{\Phi}.
    \label{eq:stabilizer_equivalence}
\end{equation}
We will refer to this intrinsic property of stabilizer codes by calling it the stabilizer equivalence.
% Should i write a short derivation for this? no right?
% Write down that acting with a stabilizer on a codedword does not change anything?


\subsection{Syndrome}\label{ssec:syndrome}

Now that we have defined the codespace, we can detect if a given eigenstate of the stabilizer group\footnote{
    Only such states are of interest, because they are the basis of our logical information.
} is in the codespace. 
For this we determine the eigenvalues of the generators of the stabilizer groups.
Based on the eigenvalues we can define an $m$-bit string, the so called syndrome $\sigma(\ket{\Psi}) \in \mathbb{Z}_2$, by
\begin{equation}
    S_i \ket{\Psi} = (-1)^{\sigma_i} \ket{\Psi}.
    \label{eq:state_syndrome}
\end{equation}
If the state is a $+1$ ($-1$) eigenstate of $i$-th generator the $i$-th bit of the syndrome takes the value $0$ ($1$).
Therefore the syndrome of the state tell us if that state belongs to the codespace, by all bits being equal $0$, or if it lies outside of it.
We will refer to the syndrome with all bits set to $0$ as the trivial syndrome.

Note that given we have $m$ generators, there are $2^m$ possible different syndromes.


We can construct a similar syndrome string for elements of the Pauli group $E\in \mathcal{P}_n$. 
Here take the element $E\in \mathcal{P}_N$ to represent an error. 
The error syndrome $\sigma(E) \in \mathbb{Z}_2$ is define by first applying the operator $E$ to any codeword $\ket{\Phi}$ and then
we can then use the definition of the syndrome of a state defined above to assign a syndrome to the error
\begin{equation}
    \sigma(E) = \sigma(E\ket{\Phi}); E \in \mathcal{P}_n; \Phi \in \mathcal{L}.
    \label{eq:error_syndrome}
\end{equation}
The resulting syndrome thus tells us how the operator $E$ would change the syndrome of a codeword. 

Looking at the definition of the $i$-th bit of the error syndrome, 
\begin{equation}
    S_i E \ket{\Phi} = (-1)^{\sigma_i} E S_i \ket{\Phi}.
\end{equation}
we can identify that the $i$-th bit of the error syndrome is defined by the commutation relation between the error and the $i$-th generator and independent of the choice of the codeword.
If they commute (anticommute) the error syndrome of the $i$-th bit has the value $0$ ($1$).
Given that both the error operator and the generator are Pauli operators, they either commute or anticommute.

Note that the choice of the specific codeword does not matter to determine the syndrome of an error operator.
The error operators changes the syndrome of each codeword by its associated error syndrome. 
Therefore it maps the whole codespace onto a different subspace where the eigenvalue for all stabilizers is no longer $+1$ but the syndrome of the error operator for each codeword.
This is a key result.

It tells us that all of those $2^m$ subspaces are isomorphic to the codespace, a fact that we are gonna use in subsection \ref{ssec:cosets_of_the_centralizer}.

Note that we can reinterpret any of those subspaces, e.g. the subspace with the syndrome $\tilde{\sigma}$, to be the codespace by redefining the stabilizer group.
For this we redefine the $i$-th generator $S_i$ to be $-S_i$ for each bit in the syndrome $\tilde{\sigma}$ which take the value $1=\tilde{\sigma}$. 
When we use the resulting set of linear independent generators to define the stabilizer group the subspace corresponding to the syndrome $\tilde{\sigma}$ using the previous set on linear independent generators, 
has now trivial syndrome assigned to it and is the codespace.
We will use this fact in the subsection about Pauli Frame Tracking again \textcolor{red}{Put ref here}.

Concluding we have $2^m$ subspaces which are isomorphic to the codespace, one for each syndrome. 
All of those subspaces are orthogonal to each other, because they have different eigenvalues/different syndromes.
Each of those subspaces is $2^k$  dimensional to encode the $k$-logical qubits, where $k=n-m$.
Therefore we can decompose the full Hilbert space $\mathcal{H}_{2^n}$ into a direct sum of the $2^m=2^{n-k}$ subspaces, each one associated with a different syndrome.


% talk a bit why we define the syndrome? motivation for decoding and such


\subsection{Centralizer, Cosets and Group Structure}\label{ssec:intro_centralizer} % find a better title 

% As eluded to in the introduction, there is a certain group structure to both the Hilbert space and Pauli group when using the stabilizer formalism which we will now investigate.
% already discovered half of this structure in the syndrome.
When looking at the syndrome we were able to find a certain structure in the Hilbert space. 
We can find some similar structure for the Pauli group $\mathcal{P}_n$, when take a look at group structure behind it. 

\subsubsection{Centralizer}

Given the observed importance of commutation relations between operators and stabilizers regarding the syndrome in the previous chapter, we are motivated to look at the centralizer\footnote{
    The centralizer is the same as the normalizer for group made out of Paulis, because Paulis either commute or anticommute. Some people refer to this group as the normalizer of the stabilizer group.
    } of the stabilizer group $\mathcal{C}(\mathcal{S})$
\begin{equation}
    \mathcal{C}(\mathcal{S}) = \left \{  P \in \mathcal{P}_n | P S_i = S_i P\  \forall S_i \in \mathcal{S}_G \right \}.
\end{equation}
The centralizer of the stabilizer group contains all the elements of the Pauli group $\mathcal{P}_n$ which commute with all elements of the stabilizer group.
To shorten the following text we will refer to the centralizer of the stabilizer group as centralizer and in case we talk about the centralizer of a different group, we will specify it.

We can deduce that all elements in the centralizer have the trivial error syndrome, because they all commute with all generators.
% Note that having a trivial syndrome is equivalent with the operators being not detectable. % TODO expand this thought


Given that the stabilizer group itself is Abelian, the centralizer contains the stabilizer group as a subgroup $\mathcal{S} \subset \mathcal{C}(S)$.
The centralizer also contains the sets $-\mathcal{S}, \pm i \mathcal{S}$. 
% I dont like the following line yet --> explain it better
But if we think about the operators in the context of possible errors on the code, or operators applied to the code, and not measurements to determine the syndrome, we can identify the phase as a global phase and are therefore not long interested in it.
Following the formalism in \cite{gottesmann_book} we will indicate the set in which we ignore the overall phases of the operator by a hat symbol, e.g.
\begin{equation}
    \hat{\mathcal{P}}_n \cong \mathcal{P}_n / \{I, iI, -I, -iI\}.
\end{equation}


Coming back to the topic at hand, we can observe that the centralizer $\hat{\mathcal{C}}(\mathcal{S})$ contains the stabilizer group $\hat{\mathcal{S}}$ 
and all other elements of the Pauli group $\hat{\mathcal{P}}_n$ which have the trivial syndrome.

Looking at the centralizer without the stabilizer $\hat{\mathcal{C}}(\mathcal{S})\setminus \hat{\mathcal{S}}$ we find an interesting set of operators.
They have the trivial error syndrome, therefore they do not change the syndrome of the state if they act on the state. 
But by definition they do not map the same codeword onto itself, otherwise they would belong to stabilizer group, which we have excluded here.
Therefore they map codeword states to other codeword states.
And we have found a set of operator which realize the logical\footnote{
    Here we refer to the logical operators which do not act trivially ($\bar{X}, \bar{Z}, \bar{Y}$). 
    The logical identity $\bar{I}$ is not changing the logical state and therefore is not changing between codewords and can be realized by any element of the stabilizer group.
} operators on the codespace.

Those operators could be applied on purpose to change from one codeword to another, which means to apply a logical operator onto the encoded qubit. 
% i do not like the following sentence
But if they happen through noise acting on the system, the logical information contained in the encoded qubit will change.
This would represent an undetectable error. 
It is undetectable because our quantum error correcting code can only tell if a state is a valid codeword. 
It can not distinguish if a state is reached by original encoding/purposefully applied logical operators and undetectable errors, hence the name. 

Therefore the set of all possible realizations of the non identity logical operators is identical to the set of undetectable errors.
% make a note that elements of the stabilizer group ignoring the phase are trivial erros ... errors not happening

% Is this useful?
Note that all operators outside of the centralizer anticommute with at least one generator of the stabilizer group and therefore all of them can be detected.
We will talk about this set of errors in subsection \ref{ssec:cosets_of_the_centralizer}.

To conclude, we detect errors by measuring the syndrome of the state. 

\subsubsection{Distance}
% Distance
At this point we can introduce an characteristic quantity of QEC codes. % more general of QEC codes
The distance $d$ which represent the minium weight\footnote{The weight of an operator is the number of qubits on which the operator acts non trivially.} that of a logical operator/undetectable error.
\begin{equation}
    d = \min_{E\in \mathcal{C}(\mathcal{S})\setminus \mathcal{S}} |E|
\end{equation}
A distance $d$ code can correct $t$ errors, with 
\begin{equation}
    t =  \left \lfloor \frac{d-1}{2} \right \rfloor.
    \label{eq:correctable_errors}
\end{equation}
The distance give us therefore an lower bound of how good our QEC code can perform. %TODO rewrite this sentence I do not like it yet!
% Explain relation between correctable error and distance? 

And now we have all necessary quantities to introduce a notation characterizing stabilizer codes.
We refer to stabilizer codes by their set of characteristic number as $[[n,k,d]]$ codes.
$n$ being the number of physical qubits, $k$ being the number of encoded qubits and $d$ is the distance of the code.


\subsubsection{Cosets of the Stabilizer Group in the Centralizer}
% I do not like this section yet, it does not feel free flowing an coherent

In subsection \ref{ssec:intro_centralizer} we introduced the centralizer. 
There we discovered that the centralizer contains the realizations of the logical operators on the encoded qubits (and the stabilizer group, 
which can be seen as the representation of the identity operator).
In this subsection we will show the underlying group theoretic structure of the centralizer, following the work of Gottesmann in \cite{gottesmann_book} (Proposition 3.10).
Understanding this structure will become especially helpful, when we discuss the decoding problem. %TODO ref the correct chapter here. 

% should I introduce the general concept of cosets here?

Central to this argumentation is that if and only if two operators from the centralizer $L_1, L_2 \in \mathcal{C}(\mathcal{S})$ act the same way on each codeword $\ket{\Phi}$, 
implying that they realize the same logical operation, 
they are in the same cosets of the stabilizer group.

To proof that we start with our assumption about the two operators
\begin{equation}
    L_1 \ket{\Phi}=L_2 \ket{\Phi}; L_1, L_2 \in \mathcal{C}(\mathcal{S}); \forall \ket{\Phi} \in \mathcal{L}
\end{equation}
which we can rearrange to 
\begin{equation}
    M \ket{\Phi} = +1\ket{\Phi}  ; \forall \ket{\Phi}\in \mathcal{L}
\end{equation}
by defining $M =  L_1^\dag L_2$.

We can see that the operator $M$ is by definition (see eq. \ref{eq:defintion_codespace_whole_group}) an element of the stabilizer group, because all codewords are in its $+1$ eigenstates.

If we rearrange the definition of $M$ we get the that $L_1$ and $L_2$ are in the same coset of the stabilizer group
\begin{equation}
    L_2 = L_1 M \text{ with } M \in \mathcal{S}. 
\end{equation}

Thus we can conclude that all the different realizations of the same logical operator are contained in the same coset of the stabilizer group.


All elements of the same coset of the stabilizer group act therefore the same on all codewords.
They all implement the same logical operation. 

% quotient group
Looking at the quotient group $\mathcal{C}(\mathcal{S})/\mathcal{S}$ we find that it contains all possible distinct logical operations\footnote{
    Note that we here keep track of the phase. One can do the same analysis for the sets in which we disregard the global phases.
} and is isomorphic to the Pauli group $\mathcal{P}_k$.
This we can write as 
\begin{equation}
    \mathcal{C}(\mathcal{S})/\mathcal{S} \cong \mathcal{P}_k.
\end{equation}
We can therefore realize logical operations on the encoded state by using an operator of the corresponding\footnote{
    We have to fix this correspondence in the beginning, see the note section following.
    } 
coset of the stabilizer group in the centralizer.
The exact choice of element from the coset is not important, all elements are logical equivalent and they only differ by an element of the stabilizer group, 
which by stabilizer equivalence (eq. \ref{eq:stabilizer_equivalence}) is to say, they do not differ. 


% Make about that the Centralizer can be written as a disjoint union of all of those cosets
Note that each cosets of the stabilizer group in the normalizer is either identical or disjoint.
Following those we can note that the centralizer $\mathcal{C}(\mathcal{S})$ can be partitioned into cosets of the stabilizer group, each realizing a unique logical operator.
For notational purposes we fix a set of operators, one for each logical operator from each cosets, and refer to the set as realizations of the logical operators $\bar{\mathcal{P}}$. 
% This set $\bar{\mathcal{P}}$ realizes each logical operator only once, and we will refer to it as the set of realizations of the logical operators.
Using it we can express the centralizer as 
\begin{equation}
    \mathcal{C}(\mathcal{S}) = \bigcup_{\bar{P} \in\bar{\mathcal{P}}} \bar{P}\mathcal{S},
    \label{eq:centralizer_partitioned_into_cosets_of_stabilizer_group}
\end{equation}
where $\bar{P}\mathcal{S}$ is the notation of the coset of the stabilizer group using the operator $\bar{P}$.

% Should I show generators of the normalizer as being generators of stabilizer group and set of realizations of the logical operators

% Should I point out that we can disregard the phase (use hat group) and the derivation still works?

% Note about freedom in assigning logical operators 
Note as well that we have some freedom in assigning the logical operators to the different cosets.
We just have to make sure that the commutations relations between the realizations of the logical operators are correct.
(As an example it is trivial to switch the logical $\bar{X}$ with the logical $\bar{Z}$ operator.)

\subsubsection{Cosets of the Centralizer}\label{ssec:cosets_of_the_centralizer}

In the previous subsection we discovered that the centralizer, which can be described as the set of all Pauli operators which have the trivial error syndrome, can be partitioned into cosets of the stabilizer group. 
Each of those cosets is associated with a logical operator on the codespace.

In subsection \ref{ssec:syndrome} we discovered that each syndrome is associated with a subspace and each of those subspaces is isomorphic to the codespace.
We also showed that we can change the syndrome of any associated space by redefining the stabilizer group.
Following the same idea we can look at all the elements of the Pauli group $\mathcal{P}_n$ and group them into sets defined by their error syndrome.

The resulting structure is that all the operators with the same error syndromes are collected in cosets of the centralizer, which therefore can be associated with this syndrome.
Similar to the section above where all operators associated to the trivial syndrome are collected in normalizer.
We can choose for each syndrome a representative error $E_{\sigma}$.

The whole Pauli group $\mathcal{P}_n$ can then be partitioned into cosets of the centralizer of the stabilizer group, where we go over all possible $2^m$ syndromes
\begin{equation}
    \mathcal{P}_n = \bigcup_{\sigma \in \{0,1\}^m} E_{\sigma} \mathcal{C}(\mathcal{S}).
\end{equation}

All of those $2^m$ cosets of the centralizer have a same inner structure.
Each of those cosets of centralizer can be partitioned into the cosets of the stabilizer group associated with a logical operator, as described in equation \ref{eq:centralizer_into_cosets_of_stabilizer_group},
where each of those cosets also inherits the representative error $E_\sigma$ of the associated syndrome.

As a result we can partition the whole Pauli group into cosets of the stabilizer group
\begin{equation}
    \mathcal{P}_n = \bigcup_{\sigma \in \{0,1\}^m} \bigcup_{\bar{P} \in\bar{\mathcal{P}}} E_{\sigma} \bar{P} \mathcal{S}.
\end{equation}

% Talk about the sizes of the different sets?


\subsection{Summary: The actual error correcting process}

So we discussed to a great detail the formalism of the stabilizer code and the underlying structure, but how does this help us to detect and correct errors?

To be able to detect errors we start in a logical codeword state.
We can verify that this state is a codeword state by measuring the generators of the stabilizer group, so by constructing the syndrome of the state.

Note that we do not care which codeword state this is, so we do not know the logical information encoded in those $k$ qubits.
And note that the logical operators by definition all commute with the stabilizers and therefore we can construct the syndrome, without collapsing any possible superposition of codeword states. 

If an error operator\footnote{
    We still assume that our error operator is an element of the Pauli group $\mathcal{P}_n$.
} $\tilde{E}$ now acts on our codeword state $\ket{\Phi}$. 
We get the resulting state $\tilde{E}\ket{\Phi}$.
Measuring the syndrome of this state, will give us the error syndrome of the error operator $\sigma(\tilde{E})$. 

This error syndrome points us directly to the coset of the centralizer with the associated syndrome $E_{\sigma(\tilde{E})}\mathcal{C}(\mathcal{S})$.
We thus know that the error $\tilde{E}$ is an element of this set $\tilde{E}\in E_{\sigma(\tilde{E})}\mathcal{C}(\mathcal{S})$.
What we do not know is from which coset of the stabilizer group this error is from.
Each coset of the stabilizer in the centralizer realizes a logical operator on the encoded qubit.
Asking the question, to which coset the error belongs is asking the question, which logical error $\bar{L}_{\tilde{E}} \in \bar{\mathcal{P}}$ does the error contain in addition to the representative error, and some element of the stabilizer group?
\begin{equation}
    \tilde{E} = E_{\sigma(\tilde{E})}\bar{L}_{\tilde{E}}\tilde{S}
\end{equation}
$\tilde{S}$ being some stabilizer element as we cant tell the exact makeup of the error.

This problem is the so called decoding problem, which is handled by so called decoders. % is solved here the correct world
We will spent some time in chapter \textcolor{red}{Put reference here} discussing this problem, as it is quite relevant to this thesis.
For now lets assume our decoder tell us from which coset of the stabilizer group the error should be from, by telling us the associated realizations of the logical operation $\bar{L}_D \in \bar{\mathcal{P}}$.

We then recover the state by choosing any element of the coset of the stabilizer group that the decoder pointed to, in the cosets of the normalizer that the syndrome pointed to.
The resulting recovery $R$ has then the form
\begin{equation}
    R = E_{\sigma(\tilde{E})}\bar{L}_{D}S_R,
\end{equation}
where we are able to identify the representative error $E_{\sigma(\tilde{E})}$ by the syndrome of the error, 
we are getting told the logical component $\bar{L}_D \in \bar{\mathcal{P}}$ by the decoder, 
and we randomly choose some element of this coset of the stabilizer group so $S_R \in \mathcal{S}$ is arbitrary.

The final question is did we recover the state correctly?
So lets look at the state $R^\dag \tilde{E} \ket{\Phi}$, 
(\textcolor{red}{either explain in detail why daggered and why it does not matter or show without, then problem with representative error!})
\begin{equation}
    R^\dag \tilde{E} \ket{\Phi} = S_R^\dag \bar{L}_D^\dag E_{\sigma(\tilde{E})}^\dag E_{\sigma(\tilde{E})}\bar{L}_{\tilde{E}}\tilde{S}\ket{\Phi}.
\end{equation}
Given that we are not interested in a conserving/recovering a global phase we can ignore those and then all those conjugates operators are equivalent to their unconjugated onces.
Applying any Pauli operator (disregarding phases) twice is identical to the identity, so $E_{\sigma(\tilde{E})}^\dag E_{\sigma(\tilde{E})} = I$, which results in 
\begin{equation}
    R^\dag \tilde{E} \ket{\Phi} = S_R \bar{L}_D \bar{L}_{\tilde{E}} \tilde{S}\ket{\Phi}.
\end{equation}
The stabilizer group element $S_R$ commutes with the logical operators.
They all belong to the centralizer as we established.
Therefore we can rewrite the expression as 
\begin{equation}
    R^\dag \tilde{E} \ket{\Phi} = \bar{L}_D \bar{L}_{\tilde{E}} S_R \tilde{S}\ket{\Phi}.
\end{equation}
By the concept of stabilizer equivalence, or just by the definition of the codeword states, the stabilizers acting on the codeword states do not change it and therefore can be absored into it.
Resulting in
\begin{equation}
    R^\dag \tilde{E} \ket{\Phi} = \bar{L}_D \bar{L}_{\tilde{E}} \ket{\Phi}.
\end{equation}

For this recovery to be successful we thus just have to choose $\bar{L}_D=\bar{L}_{\tilde{E}}$. 
In this case we would recover our original codeword state $\ket{\Phi}$.
If they are not identical, they would implement a different logical operator $\bar{L}_\Delta = \bar{L}_D \bar{L}_{\tilde{E}}$
, as indicated by the isomorphism between the choosen realizations of the logical operators $\bar{\mathcal{P}}$ and the Pauli group for the the encoded logical qubit $\mathcal{P}_k$.
Which then would change our codeword resulting in an unsuccessful recovery.






% then discuss all errors just beeing pauli, and then that the error itself are not identifable we can just identify the coset of the normalizer
% task of the decoder is then finding the correct pauli/coset of pauli
% here bring into play stabilizer equivalence when it comes to corrections 

% lastly introduce the distance of the code and set it into relation with the amount of correctable errors.

\section{Kitaev Surface Code}\label{sec:kitaev_surface_code}

% defintion and properties of kitaev surface code
% stabilizer and logical operators 
% scaling of such a code
% connection to toric code


% \subsubsection{Logical State Preparation}

\section{Threshold Theorem}

% talk about why scale anyways 
% at which point encoding a qubit makes actual sense -> pseudo threshold
% threshold of a code vs distance of a code


\section{Pauli Errors}\label{sec:pauli_errors}

% 1. realworld physical noise process -> model those in theory as pauli errors <- this is an underlying assumption
% endless physical errors (they are continous)

% 1. talk about digitization of errors 
% name knill laflamme QEC condition, 

% any error can be approximated/expanded (which of both?) in discrete pauli basis
% syndrome measurements projects into those error components with certain probability 
% we just need to correct the projected error by applying a pauli 

% Discuss the pauli group here a bit
% all paulis either commute or anticommute
% generators of P_n
% weight of an error

% markovian 
% we do not simulate erasure and leakage errors


We will discuss the exact shape of the noise model used in this thesis in the following, but we can already state that we assume Markovian noise, which means that our noise is not correlated in time. 


% limit ourselfs to computational errors and the following channels
% 2. introduce phase-,bit-flip error and depolirizing noise 
% 3. Further talk about structure and sources of noise -> generalise in models which are split up as follows

\subsection{Code Capacity Noise Model}

% 0. Just errors on data qubits 
% 1. Simplest model -> assumes noise free QEC <- somewhat faulty assumption -> leads to next error model

\subsection{Circuit Level Noise Model}

% 0. Ancilla qubits and gates are faulty 
%     + more realworld like
%     + motivate the need to think about syndrome extraction
% 0. explain concept of fault tolerance and residual error
%     + error on qubits only with prob p
%     + introduce the need for fault tolerance check
% 2. Add two qubit depolirizing noise, due to CNOT gates
% 3. describe the resulting full error model/show graph


\subsubsection{Phenomological Noise Model}

% Do I actually describe the correct model here?
% -> I want to use this later as a references when I want to motivate the propagated error model

% 0. Definition 
% 1. Why useful?


\section{Fault Tolerance}



